---
title: "Experimental Summarization Notebook"
output: html_document
params:
  snapshot_name: "official3"
  snapshot_path: "/data/COVID-Twitter/analysis/snapshots"
---

```{r setup, include=FALSE}
# Required R package installation:
# These will install packages if they are not already installed
# Set the correct default repository
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

if (!require("knitr")) {
  install.packages("knitr")
  library(knitr)
}

if (!require("dplyr")) {
  install.packages("dplyr")
  library(dplyr)
}

if (!require("kableExtra")) {
  install.packages("kableExtra")
  library(kableExtra)
}

knitr::opts_chunk$set(echo = TRUE)

unlockBinding("params", env = .GlobalEnv)
snapshot_filename <- file.path(params$snapshot_path, paste(params$snapshot_name, ".Rdata", sep=""))
load(snapshot_filename)

source("Summarizer.R")
source("text_helpers.R")
```

## Summarization Strategies:
Each strategy determines how N tweets are selected from each cluster and subcluster to assemble the "document"
that is given to the summarization model.

### Centers
Cluster Summaries:    N nearest neighbors to cluster center are selected.
Subcluster Summaries: N nearest neighbors to subcluster center are selected.

### Subcenters
Cluster Summaries:    M nearest neighbors to each subcluster center are selected, where sum(M) <= N.
                      for example for N=20, a cluster with 15 subclusters would select the single nearest
                      neighbor to each subcluster center.
Subcluster Summaries: N nearest neighbors to subcluster center are selected.

### Heirarchy Centers
Cluster Summaries:    All the subcluster summaries within this cluster are used in lieu of tweets.
Subcluster Summaries: N nearest neighbors to subcluster center are selected.


## Summaries:

```{r, echo=FALSE}
summarize_max_len <- 60
summarize_num_beams <- 6
summarize_temperature <- 1.0

experiments <- expand.grid(model=c("distilbart-xsum-12-6", 
                                   "distilbart-cnn-12-6"),
                           strategy=c("centers", 
                                      "subcenters", 
                                      "heirarchy_centers"),
                           stringsAsFactors=FALSE)

results <- list()
for (i in 1:nrow(experiments)) {
  print(paste0("Runnning experiment ", i, " of ", nrow(experiments), "..."))
  
  model <- experiments[i, "model"]
  model_name <- paste0("sshleifer/", model)
  strategy <- experiments[i, "strategy"]
  
  summaries.df <- tweet.vectors.df[tweet.vectors.df$vector_type != "tweet", 
                                   c("vector_type", "cluster", "subcluster")]
  
  summaries.df$model <- model
  summaries.df$strategy <- strategy
  
  #compute the subcluster text for summarization
  subcluster_summaries.df <- summaries.df[summaries.df$vector_type == "subcluster_center",]
  subcluster_summaries.df$text_for_summary <- mapply(function(cluster, subcluster) {
    nearest_center <- clusters[[cluster]]$subclusters[[subcluster]]$nearest_center
    return(concat_text_for_summary(nearest_center, summarize_center_nn))
  }, subcluster_summaries.df$cluster, subcluster_summaries.df$subcluster)
  summaries.df[rownames(subcluster_summaries.df), "text_for_summary"] <- subcluster_summaries.df$text_for_summary
  
  #do the summarization for all subclusters in a single batch
  subcluster_summaries.df$summary <- summarize(text=subcluster_summaries.df$text_for_summary,
                                               max_len=summarize_max_len,
                                               num_beams=summarize_num_beams,
                                               temperature=summarize_temperature,
                                               model=model_name,
                                               summarizer_url=summarizer_url)
  summaries.df[rownames(subcluster_summaries.df), "summary"] <- subcluster_summaries.df$summary
  
  
  #compute the cluster text for summarization
  cluster_summaries.df <- summaries.df[summaries.df$vector_type == "cluster_center",]
  cluster_summaries.df$text_for_summary <- sapply(cluster_summaries.df$cluster, function(cluster) {
      if (strategy == "subcenters") {
        subcenter_nn <- max(1, floor(summarize_center_nn/cluster.k))
        nearest_center <- data.frame(full_text=c(sapply(1:cluster.k, function(subcluster) {
          clusters[[cluster]]$subclusters[[subcluster]]$nearest_center[1:subcenter_nn, "full_text"]
        })))
      } else if (strategy == "heirarchy_centers") {
        nearest_center <- data.frame(
          full_text=subcluster_summaries.df[subcluster_summaries.df$cluster==cluster, "summary"])
      } else {
        nearest_center <- clusters[[cluster]]$nearest_center
      }
      return(concat_text_for_summary(nearest_center, summarize_center_nn))
  })
  summaries.df[rownames(cluster_summaries.df), "text_for_summary"] <- cluster_summaries.df$text_for_summary
  
  #do the summarization for all clusters in a single batch
  cluster_summaries.df$summary <- summarize(text=cluster_summaries.df$text_for_summary,
                                            max_len=summarize_max_len,
                                            num_beams=summarize_num_beams,
                                            temperature=summarize_temperature,
                                            model=model_name,
                                            summarizer_url=summarizer_url)
  summaries.df[rownames(cluster_summaries.df), "summary"] <- cluster_summaries.df$summary

  
  results[[i]] <- summaries.df
}

summaries.df <- bind_rows(results)

summaries.df <- summaries.df %>% 
  filter(vector_type == "cluster_center" | strategy == "centers") %>%
  dplyr::arrange(vector_type, cluster, subcluster, model, strategy)
summaries.df[summaries.df$vector_type == "cluster_center", "subcluster"] <- NA
  
saveRDS(summaries.df, "summaries.df.rds")
```

```{r, echo=FALSE}
kable(summaries.df[,setdiff(colnames(summaries.df), "text_for_summary")]) %>% kable_styling(font_size=11)
```